Weighted Model Counting with Conditional Weights for Bayesian Networks

### 2021-02-09 doc/paper2/paper.tex
Related work section

\section{Introduction}
% TODO: introduce \textsf{DPMC} \citep{DBLP:conf/cp/DudekPV20}

@@

\section{Related Work} \label{sec:related}
% TODO: tackle a comparable problem, but their encoding was never implemented or
% empirically tested (i.e., the implementation is left as future work)
% \citep{DBLP:conf/scai/VomlelT13}

% TODO: introduce \texttt{bklm16} \citep{DBLP:conf/ecai/BartKLM16}

Hitherto, four techniques have been proposed for encoding Bayesian networks into
instances of WMC. We will identify them based on the initials of authors as well
as publications years: \texttt{d02} \citep{DBLP:conf/kr/Darwiche02},
\texttt{sbk05} \citep{DBLP:conf/aaai/SangBK05}, \texttt{cd05}
\citep{DBLP:conf/ijcai/ChaviraD05}, and \texttt{cd06}
\citep{DBLP:conf/sat/ChaviraD06}. Below we summarise the observed performance
differences among them. \citet{DBLP:conf/aaai/SangBK05} claim that
\texttt{sbk05} is a smaller encoding than \texttt{d02} with respect to both the
number of clauses and the number of variables but provide no experimental
comparison. \citet{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with
\texttt{d02} by measuring the time it takes to compile either encoding into an
arithmetic circuit (but do not measure inference time). The more recent encoding
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\citep{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \citep{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with \textsf{Cachet} \citep{DBLP:conf/sat/SangBBKP04} with
the compile times of \texttt{cd05} and \texttt{cd06}, but only on five (types
of) instances. In these experiments, \texttt{cd06} is always faster than
\texttt{cd05}, while the comparison with \texttt{sbk05} is mixed. Sometimes
\texttt{cd06} is orders of magnitude faster than \texttt{sbk05}, sometimes
slightly slower. The performance difference between \texttt{sbk05} and
\texttt{cd05} is even harder to judge: \texttt{sbk05} is better on three out of
five instances and worse on the remaining two. Based on this description, one
would expect \texttt{cd06} to be faster than both \texttt{cd05} and
\texttt{sbk05}, both of which should be faster than \texttt{d02}. The
experiments in \cref{sec:experiments}, however, strongly disagree with this
prediction, showing that the quality of an encoding depends strongly on the
underlying search algorithm or compilation technique.

@@

\section{Experimental Comparison} \label{sec:experiments}
% TODO: introduce \textsf{query-dnnf}
% \footnote{\url{http://www.cril.univ-artois.fr/kc/d-DNNF-reasoner.html}} (also
% talk more about how Cachet and Ace are used)
% TODO: included original WMC algorithms
% TODO: Even instances that are solved in less than 0.1 s are too big for edge
% count to be efficiently calculated.
% TODO: my plot's sbk05 data should match (and matches) the ADDMC/DPMC vs c2d
% comparison in the paper (although my sbk05 data only roughly corresponds to
% half of their dataset)
% TODO: mention that both encoding and inference times are limited to 1000 s
% TODO: mention that we don't run ADDMC+cd05 and ADDMC+cd06 because they don't produce correct answers

@@

\begin{table}
  \centering
  \caption{The numbers of instances (out of 1466) solved by
    each combination of algorithm and encoding (uniquely, faster than others, and in total).}%
  \begin{tabular}{lrrr}
    \toprule
    Algorithm \& Encoding & Unique & Fastest & Total \\
    \midrule
    $\textsf{Ace} + \texttt{cd05}$ & \textcolor{gray}{0} & 55 & 1169 \\
    $\textsf{Ace} + \texttt{cd06}$ & \textcolor{red}{34} & 218 & \textcolor{red}{1259} \\
    $\textsf{Ace} + \texttt{d02}$ & \textcolor{gray}{0} & 46 & 993 \\
    $\textsf{ADDMC} + \texttt{bklm16}$ & \textcolor{gray}{0} & 29 & \textcolor{blue}{617} \\
    $\textsf{ADDMC} + \texttt{cw}$ & 14 & \textcolor{red}{770} & 919 \\
    $\textsf{ADDMC} + \texttt{d02}$ & \textcolor{gray}{0} & \textcolor{gray}{0} & 703 \\
    $\textsf{ADDMC} + \texttt{sbk05}$ & \textcolor{gray}{0} & \textcolor{gray}{0} & 729 \\
    $\textsf{Cachet} + \texttt{sbk05}$ & 13 & 229 & 928 \\
    $\textsf{query-dnnf} + \texttt{bklm16}$ & \textcolor{gray}{0} & 3 & 1017 \\
    \bottomrule
  \end{tabular}
  \label{tbl:tallies}
\end{table}

@@

\begin{figure*}[t]
  \centering
  \input{melt.tex}%
  \caption{TODO: describe this in the text}
  \label{fig:melted}
\end{figure*}

@@

For all encodings other than \texttt{cw}, we use their implementation in
\textsf{Ace}~3.0\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} with
\texttt{-encodeOnly} and \texttt{-noEclause} flags. However, \textsf{Ace} was
not used to encode evidence, as preliminary experiments revealed that the
evidence-encoding implementation contains bugs that can lead to incorrect
answers or a Java exception being thrown on some instances of the data set (and
the source code is not publicly available). Instead, we simply list all the
evidence as additional clauses in the encoding.

@@

\begin{table} % TODO: add bklm16
  \centering
  \caption{Asymptotic upper bounds on the numbers of variables and clauses/ADDs
    for each encoding.}
  \label{tbl:asymptotes}
  \begin{tabular}{lcc}
    \toprule
    Encoding(s) & Variables & Clauses/ADDs \\
    \midrule
    \texttt{cd05}, \texttt{cd06}, \texttt{sbk05} & $O(nv^{d+1})$ & $O(nv^{d+1})$ \\
    \texttt{cw} & $O(nv)$ & $O(nv^2)$ \\
    \texttt{d02} & $O(nv^{d+1})$ & $O(ndv^{d+1})$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\section{Discussion and Conclusion}
% TODO: mention that DPMC is the next step

>>>>

\section{Introduction}

@@

\section{Related Work} \label{sec:related}

Performing inference on Bayesian networks by encoding them into instances of WMC
is a well-established idea with a history of almost twenty years. Five encodings
have been proposed so far (we will identify them based on the initials of
authors as well as publications years): \texttt{d02}
\citep{DBLP:conf/kr/Darwiche02}, \texttt{sbk05} \citep{DBLP:conf/aaai/SangBK05},
\texttt{cd05} \citep{DBLP:conf/ijcai/ChaviraD05}, \texttt{cd06}
\citep{DBLP:conf/sat/ChaviraD06}, and \texttt{bklm16}
\citep{DBLP:conf/ecai/BartKLM16}\footnote{\citet{DBLP:conf/scai/VomlelT13} also
  propose an encoding, but only for networks of a particular bipartite structure
  and without any evaluation.}. Below we summarise the observed performance
differences among them.

\citet{DBLP:conf/aaai/SangBK05} claim that
\texttt{sbk05} is a smaller encoding than \texttt{d02} with respect to both the
number of clauses and the number of variables but provide no experimental
comparison. \citet{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with
\texttt{d02} by measuring the time it takes to compile either encoding into an
arithmetic circuit. They show that
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\citep{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \citep{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with \textsf{Cachet} \citep{DBLP:conf/sat/SangBBKP04} with
the compile times of \texttt{cd05} and \texttt{cd06}, but only on five (types
of) instances. In these experiments, \texttt{cd06} is always faster than
\texttt{cd05}, while the comparison with \texttt{sbk05} is mixed. Sometimes
\texttt{cd06} is orders of magnitude faster than \texttt{sbk05}, sometimes
slightly slower. The performance difference between \texttt{sbk05} and
\texttt{cd05} is even harder to judge: \texttt{sbk05} is better on three out of
five instances and worse on the remaining two. Finally,
\citet{DBLP:conf/ecai/BartKLM16} introduce \texttt{bklm16} and show that it has
both fewer variables and fewer clauses than \texttt{cd06}. Their experiments
show \texttt{bklm16} to be superior to \texttt{cd06} with respect to both
compilation time and encoding size when both are compiled using \textsf{c2d}
\citep{DBLP:conf/ecai/Darwiche04} but inferior to \texttt{cd06} when
\texttt{cd06} is compiled using
\textsf{Ace}\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} (which still uses
\textsf{c2d} but considers the structure of the Bayesian network along with its
encoding).

Our experiments in \cref{sec:experiments} compare the five previous encodings
along with a newly proposed one---with originally-proposed WMC algorithms as
well as with \textsf{ADDMC}---with respect to the total time it takes to encode
a Bayesian network, compile it or run a WMC algorithm on it, and extract the
(numerical) answer. We confirm some of the findings outlined in this section
while also observing that the (relative and absolute) performance of each
encoding depends on the WMC algorithm in use, and smaller encodings are not
necessarily faster.

@@

\section{Experimental Comparison} \label{sec:experiments}
% TODO: introduce \textsf{query-dnnf}
% \footnote{\url{http://www.cril.univ-artois.fr/kc/d-DNNF-reasoner.html}} (also
% talk more about how Cachet and Ace are used)
% TODO: included original WMC algorithms
% TODO: Even instances that are solved in less than 0.1 s are too big for edge
% count to be efficiently calculated.
% TODO: my plot's sbk05 data should match (and matches) the ADDMC/DPMC vs c2d
% comparison in the paper (although my sbk05 data only roughly corresponds to
% half of their dataset)
% TODO: mention that both encoding and inference times are limited to 1000 s
% TODO: mention that we don't run ADDMC+cd05 and ADDMC+cd06 because they don't produce correct answers
% TODO: the -implicit flag is disabled for good reasons for bklm16
% TODO: potential appendix: exact arguments to each piece of software

@@

\begin{table}
  \centering
  \caption{The numbers of instances (out of 1466) solved by
    each combination of algorithm and encoding (uniquely, faster than others, and in total).}%
  \begin{tabular}{lrrr}
    \toprule
    Algorithm \& Encoding & Unique & Fastest & Total \\
    \midrule
    $\textsf{Ace} + \texttt{cd05}$ & \textcolor{gray}{0} & 55 & 1169 \\
    $\textsf{Ace} + \texttt{cd06}$ & \textbf{34} & 218 & \textbf{1259} \\
    $\textsf{Ace} + \texttt{d02}$ & \textcolor{gray}{0} & 46 & 993 \\
    $\textsf{ADDMC} + \texttt{bklm16}$ & \textcolor{gray}{0} & 29 & \textcolor{gray}{617} \\
    $\textsf{ADDMC} + \texttt{cw}$ & 14 & \textbf{770} & 919 \\
    $\textsf{ADDMC} + \texttt{d02}$ & \textcolor{gray}{0} & \textcolor{gray}{0} & 703 \\
    $\textsf{ADDMC} + \texttt{sbk05}$ & \textcolor{gray}{0} & \textcolor{gray}{0} & 729 \\
    $\textsf{Cachet} + \texttt{sbk05}$ & 13 & 229 & 928 \\
    $\textsf{query-dnnf} + \texttt{bklm16}$ & \textcolor{gray}{0} & 3 & 1017 \\
    \bottomrule
  \end{tabular}
  \label{tbl:tallies}
\end{table}

@@

%\begin{figure*}[t] % TODO: describe this in the text
%  \centering
%  \input{melt.tex}%
%  \caption{Average encoding and inference time of each encoding with their
%    original WMC algorithm (on the left) and with \textsf{ADDMC} (on the right)
%    on \emph{solved} instances. The error bars denote \SI{10}{\percent} of
%    standard deviation.}
%  \label{fig:melted}
%\end{figure*}

@@

For all encodings other than \texttt{cw}, we use their implementation in
\textsf{Ace}~3.0 with
\texttt{-encodeOnly} and \texttt{-noEclause} flags. However, \textsf{Ace} was
not used to encode evidence, as preliminary experiments revealed that the
evidence-encoding implementation contains bugs that can lead to incorrect
answers or a Java exception being thrown on some instances of the data set (and
the source code is not publicly available). Instead, we simply list all the
evidence as additional clauses in the encoding.

@@

\begin{table}
  \centering
  \caption{Asymptotic upper bounds on the numbers of variables and clauses/ADDs
    for each encoding.}
  \label{tbl:asymptotes}
  \begin{tabular}{m{0.4\linewidth} c c}
    \toprule
    Encoding(s) & Variables & Clauses/ADDs \\
    \midrule
    \texttt{bklm16}, \texttt{cd05}, \texttt{cd06}, \texttt{sbk05} & $O(nv^{d+1})$ & $O(nv^{d+1})$ \\
    \texttt{cw} & $O(nv)$ & $O(nv^2)$ \\
    \texttt{d02} & $O(nv^{d+1})$ & $O(ndv^{d+1})$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\section{Discussion and Conclusion}
% TODO: mention that DPMC is the next step
% TODO: which algorithm comes out on top depends on the data set
% TODO: emphasise that BN encodings have never been compared w.r.t. time before
